---
title: "demonstration"
author: "Thijs Janzen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example: abc}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ltables

Ltables describe the procession of new species arising through speciation,
where each species is tracked in the table on a separate row, indicating the
time of birth, parent ID, self ID and finally the time of death. If the species
did not die (go extinct), the time of death is -1. 
We will demonstrate ltables using the included simulation code of the
birth-death model:

```{r yule}
res <- ltable::sim_bd(parameters = c(0.5, 0),
                      crown_age = 5,
                      seed = 4)
res$ltable
```

Ltables can be converted to phy objects using either `[DDD::L2phylo]` or 
`[treestats::l_to_phylo]`. However, by this conversion, some information is lost,
primarily the information on parentage. Instead, this R package allows for the 
direct visualisation of the Ltable, leaving information of parentage available.
```{r plot}
ltable::plot_ltable(res$ltable)
```
Compare that to the resulting phylogeny:
```{r plot_ape}
plot(res$phy, direction = "up")
```

By default, branches are colored using a continuous gradient - this is 
especially useful for large trees with many species in it. Instead, we can also
color the tree giving distinct colors to the branches:
```{r plot2}
ltable::plot_ltable(res$ltable, use_factor_ID = TRUE)
```

### Extinct branches
Ltables can also have extinct species, and those can be plotted as well:
```{r bd}
res <- ltable::sim_bd(parameters = c(0.5, 0.2),
                      crown_age = 5,
                      seed = 4)
res$ltable
ltable::plot_ltable(res$ltable, use_factor_ID = TRUE)
```

### Using DDD
The DDD package by default generates ltables, we can use those as well:
```{r ddd}
res <- DDD::dd_sim(pars = c(0.5, 0.1, 100),
                   age = 5,
                   ddmodel = 1)
ltable::plot_ltable(res$L, use_factor_ID = TRUE)
```
